Can Auxiliary Indicators Improve COVID-19 Forecasting and Hotspot Prediction?

### 2021-10-31 forecast/paper/supplement-text.tex
Intermediate progress on supp text

\section{Aggregation with Geometric Mean}

In this section, we consider using the geometric mean instead of the arithmetic
mean when aggregating the weighted interval score (WIS) across location-time
pairs.  There are three reasons why using the geometric mean may be desirable.
\begin{enumerate}
\item WIS is right-skewed, being bounded below by zero and having occasional 
  very large values.  Figure \ref{fig:wis-densities} illustrates that the
  densities appear roughly log-Gaussian.  The geometric mean is a natural choice  
  in such a context since the relative ordering of forecasters is determined by
  the arithmetic mean of the {\em logarithm} of their WIS values.
\item In the main paper, we report the ratio of the mean WIS of a forecaster to
  the mean WIS of the baseline forecaster. Another choice could be to take the
  mean of the ratio of WIS values for the two methods. This latter choice would
  penalize a method less for doing poorly where the baseline forecaster also
  does poorly.  Using instead the geometric mean makes the order of aggregation
  and scaling immaterial since the ratio of geometric means is the same as the 
  geometric mean of ratios.
\item If one imagines that a forecaster's WIS is composed of multiplicative
  space-time effects $S_{\ell,t}$ shared across all forecasters,
  i.e., $\WIS(F_{\ell,t,f},Y_{\ell,t})=S_{\ell,t}E_{f,t}$ with $E_{f,t}$ a
  forecaster-specific error, then taking the ratio of two forecasters' geometric
  mean WIS values will effectively cancel these space-time effects. 
\end{enumerate}

Figure \ref{fig:fcast-adjusted} uses the geometric mean for aggregation.
Comparing this with Figure 3 of the main paper, we see that the main conclusions
are largely unchanged; however, \chngcli~now appears better than \ar.  This
behavior would be expected if \chngcli's poor performance is attributable to a
relatively small number of large errors (as opposed to a large number of
moderate errors).  Indeed, Figure 5 of the main paper further corroborates this,
in which we see the heaviest left tails occuring for \chngcli.

\section{Comparing with the COVID-19 Forecast Hub}

Since July of 2020, teams have been submitting real time forecasts
to the COVID-19 Forecast Hub \cite{ForecastHub}. While a number of forecast
targets can be submitted, the most related to that used in this paper---the
7-day trailing average of COVID-19 case incidence rates at the Hospital Referral
Region---are state-level forecasts for total case incidence over an epiweek.
Essentially, for each state, teams are asked to predict the total number of
cases that will be recorded over the week from Monday to Sunday. Forecasts must
be submitted by Tuesday for the current Sunday as well as the next three
Sundays. This corresponds to producing forecasts for 5, 12, 19, and 26 days
ahead. Some teams choose to submit their forecasts a few days earlier, but
always for the same targets.

To properly compare the models presented in this paper with those submitted to
the Hub, we limit ourselves to the same set of forecast dates and states, and
we predict case incidence over a week by: (1) predicting the 7-day trailing
average case rate (as in the manuscript);
(2) multiplying by 7 to give the total over a week; (3) multiplying by the state
population as reported in the 2019 US Census Population Estimate (available in
the Covidcast R package \cite{CovidcastR}). Finally, we perform the same
evaluations as in the manuscript Figure 3 and also using geometric mean as
described above.
For the purpose of the comparison, we show only the AR model along with
submissions collected by the Forecast Hub in Figure \ref{fig:compare-to-hub}.
Since not all teams submitted forecasts for the entire period, we
only display those which submitted for at least 6 weeks. By both metrics,
the AR model examined in this paper is competitive with the submitted by top teams, even
beating the Ensemble for smaller target horizons.

>>>>

\section{Aggregation With Geometric Mean}

In this section, we consider using the geometric mean instead of the usual
(arithmetic) mean when aggregating the weighted interval score (WIS) across
location-time pairs.  There are two reasons why using the geometric mean may be
desirable.

\begin{enumerate}
\item WIS is right-skewed, being bounded below by zero and having occasional 
  very large values.  Figure \ref{fig:wis-densities} illustrates that the
  densities appear roughly log-Gaussian.  The geometric mean is a natural choice   
  in such a context since it can be viewed as a measure of centrality on the log
  scale (it is the exponential of the arithmetic mean of log values).
\item In the main paper, we report the ratio of the mean WIS of a forecaster to
  the mean WIS of the baseline forecaster. Another choice could be to take the
  mean of the ratio of WIS values for the two methods. This latter choice would
  penalize a method less for doing poorly where the baseline forecaster also
  does poorly.\footnote{In a sense, this is implicitly estimating a
    nonparametric space-time effect for forecaster error, and assuming that has
    a shared, multiplicative contribution to forecaster errors.  That is, if one
    imagines that a forecaster's WIS is composed of multiplicative space-time
    effects $S_{\ell,t}$ shared across all forecasters,
    $\WIS(F_{\ell,t,f},Y_{\ell,t}) = S_{\ell,t} E_{f,t}$ with $E_{f,t}$ a 
    forecaster-specific error, then taking the ratio of individual WIS values
    cancels these space-time effects.}  Using instead the geometric
  mean makes the order of aggregation and scaling immaterial since the ratio of
  geometric means is the same as the geometric mean of ratios.
\end{enumerate}

Figure \ref{fig:fcast-adjusted} uses the geometric mean for aggregation.
Comparing this with Figure 3 in the main paper, we see that the main conclusions
are largely unchanged; however, \chngcli~now appears better than \ar.  This
behavior would be expected if \chngcli's poor performance is attributable to a
relatively small number of large errors (as opposed to a large number of
moderate errors).  Indeed, Figure 5 of the main paper further corroborates this,
in which we see the heaviest left tails occur for \chngcli.

\section{Comparing With the COVID-19 Forecast Hub}

Since July of 2020, modelers have been submitting real-time forecasts of
COVID-19 case incidence to the COVID-19 Forecast Hub \cite{ForecastHub}. This 
(along with forecasts of hospitalizations and deaths collected in the same Hub)
serves as the source of the CDC's official communications on COVID forecasting.

Our goal in this section is to compare the AR model and indicator models to
those in the Hub, in terms of forecast errors aggregated over the same forecast
tasks, to give a clear sense for how robust and effective the models we choose
to investigate in the paper are relative to those in common operational use.
This was prompted by a question from an anonymous reviewer of this paper, who
asked why we chose to build our analysis of indicator utility around the AR
model in the first place, and why we did not investigate others (say, the SIR
model or more complex mechanistic models of disease transmission) that have
occupied more of the spotlight over the course of the pandemic.  The analysis
presented here corroborates the claim that an AR model is simple, and yet,
properly trained---using a quantile loss to directly estimate multiple
conditional quantiles, a trailing training window of 21 days, pooling across all
locations jointly, and fitting to case rates rather than counts---all as we do
in the models in the main paper, it can be robust and effective, performing 
competitively with the top models from COVID-19 Forecast Hub.

The closest forecast target in the Hub to that used in this paper are
state-level case incidence over an epiweek---defined by the sum of new case
counts reported between a Sunday and the following Saturday (inclusive). Our
target is 7-day trailing average of COVID-19 case incidence rates at the HRR
level, which is different in three regards:

\begin{enumerate}
\item temporal resolution (daily versus weekly);
\item geographic resolution (HRRs versus states); and 
\item scale (rates versus counts).
\end{enumerate}

While the first and third of these differences can be easily addressed post 
hoc---meaning, we can take always take our model's outputs and multiply them by
7 in order to track the incidence over any given trailing week, and rescale them
per-location by population to put them on the count scale---the second
difference is not easy to adjust post hoc, because a quantile is not linear
operator (a quantile of a linear combination of random variables is not simply
the linear combination of their quantiles, but rather, depends intricately on
the correlations between random variables).

Therefore, to make a comparison to models in the Hub in as direct a way as
possible, we retrained our models over the same forecast period as in the main
paper, and with the same general setup entirely, except at the state rather than 
HRR level.  We then rescaled them post hoc to account for the different temporal
resolution and the rate versus count scale (first and third items on the above
list).  The results are given in Figure \ref{fig:compare-to-hub}.  The
evaluation was carried out exactly as in the main paper, and the figure displays
both mean WIS and geometric mean WIS, as a function of ahead.  Furthermore, to
account for missingness (not all teams submitted forecasts to the Hub for all
locations and ahead values for the entire period), we first dropped any
forecaster than submitted for less than 6 weeks, and then restricted the
aggregation metrics (mean or geometric mean) to errors from commonly available
forecast tasks.

By either metric, mean and geometric mean WIS, relative to baseline, we can see
in Figure \ref{fig:compare-to-hub} that the AR model examined in this paper is 
competitive with top models in the Hub, even outperforming the Forecast Hub
ensemble model for smaller ahead values.  The same general conclusion can be
drawn for the indicators models as well.  However, interestingly, a close
inspection reveals that the AR model here is for the most part in the ``middle
of the pack'' when compared to the indicator models, and only the Google-AA
model offers clear improvement over AR for all aheads.  This is likely due to
the fact that at the state level, the signal-to-noise ratio (SNR) is generally
higher, and AR model provides a higher standard (on which to expect improvement
using an auxiliary indicator).  At the HRR level, with lower SNR, using
indicators as simple additional linear features in the AR model provides an
important enough variance reduction

% Essentially, for each state, teams are asked to predict the total number of
% cases that will be recorded over the week from Monday to Sunday. Forecasts must
% be submitted by Tuesday for the current Sunday as well as the next three
% Sundays. This corresponds to producing forecasts for 5, 12, 19, and 26 days
% ahead. Some teams choose to submit their forecasts a few days earlier, but
% always for the same targets.

To properly compare the models presented in this paper with those submitted to
the Hub, we limit ourselves to the same set of forecast dates and states, and
we predict case incidence over a week by: (1) predicting the 7-day trailing
average case rate (as in the manuscript);
(2) multiplying by 7 to give the total over a week; (3) multiplying by the state
population as reported in the 2019 US Census Population Estimate (available in
the Covidcast R package \cite{CovidcastR}). Finally, we perform the same
evaluations as in the manuscript Figure 3 and also using geometric mean as
described above.
For the purpose of the comparison, we show only the AR model along with
submissions collected by the Forecast Hub in Figure \ref{fig:compare-to-hub}.
Since not all teams submitted forecasts for the entire period, we
only display those which submitted for at least 6 weeks.
